---
title: "19130 CAR product cluster annotation"
author: "heinin"
date: "2024-11-11"
output:
  workflowr::wflow_html:
    code_folding: hide
  html_notebook: default
editor_options:
  chunk_output_type: console
---

## T cell annotations

### Libraries and environment variables

```{r, message = F, warning = F}

library(workflowr)
library(Seurat)
library(googlesheets4)
library(tidyverse)
library(plyr)
library(UpSetR)
library(scImmuCC)

setwd("/home/hnatri/19130_CAR_production/")
set.seed(1234)
options(future.globals.maxSize = 30000 * 1024^2)
reduction <- "integratedSCTumap"

source("/home/hnatri/19130_CAR_production/code/CART_plot_functions.R")
source("/home/hnatri/19130_CAR_production/code/colors_themes.R")

# Cluster annotations
gs4_deauth()
cluster_annot  <- gs4_get("https://docs.google.com/spreadsheets/d/1DstdK4vXjk7ZEeBalW_4CA6Vmd57BGENUx3iMDRLszk/edit?usp=sharing")
sheet_names(cluster_annot)
cluster_annot <- read_sheet(cluster_annot, sheet = "Cluster annotations")
head(cluster_annot)

# Canonical T cell markers
gs4_deauth()
canonical_markers  <- gs4_get("https://docs.google.com/spreadsheets/d/186PIcU3Jb0Xm3IuZgENEChluykvwsK4JnHF8FbeJASA/edit?usp=sharing")
sheet_names(canonical_markers)
canonical_markers <- read_sheet(canonical_markers, sheet = "T cells, gene sets")
head(canonical_markers)

```

### Import data

```{r, message = F, warning = F, fig.width = 8, fig.height = 4}

# Integrated data
integrated_seurat <- readRDS("/tgen_labs/banovich/pediatric_CAR-T/02_production/19130_CAR_production_integrated_ProjecTILs.rds")
integrated_seurat$cluster <- integrated_seurat$integratedSCTsnn_res.1
integrated_seurat <- JoinLayers(integrated_seurat, assay = "RNA")
integrated_seurat <- NormalizeData(integrated_seurat)

DimPlot(integrated_seurat,
        group.by = "cluster",
        cols = product_cluster_col,
        reduction = reduction,
        label = T,
        label.box = T,
        label.size = 3,
        repel = T,
        raster = T,
        raster.dpi = c(1024, 1024),
        pt.size = 3) +
  ggtitle("integratedSCTsnn_res.1") +
  theme_classic() +
  NoLegend() +
  NoAxes() +
  coord_fixed(1)

DimPlot(integrated_seurat,
        group.by = "functional.cluster",
        reduction = reduction,
        raster = T,
        raster.dpi = c(1024, 1024),
        pt.size = 3) +
  ggtitle("ProjecTILs functional.cluster") +
  theme_classic() +
  NoAxes() +
  coord_fixed(1)

```

```{r, message = F, warning = F, fig.width = 5, fig.height = 5}

DimPlot(integrated_seurat,
        group.by = "Phase",
        reduction = reduction,
        raster = T,
        raster.dpi = c(1024, 1024),
        pt.size = 3) +
  ggtitle("Phase") +
  theme_classic() +
  NoAxes() +
  coord_fixed(1)

```


```{r, message = F, warning = F, fig.width = 8, fig.height = 16}

DimPlot(integrated_seurat,
        group.by = "cluster",
        split.by = "cluster",
        cols = product_cluster_col,
        reduction = reduction,
        ncol = 5,
        label = T,
        label.box = T,
        label.size = 3,
        repel = T,
        raster = T,
        raster.dpi = c(1024, 1024),
        pt.size = 3) +
  ggtitle("integratedSCTsnn_res.0.8") +
  theme_classic() +
  NoLegend() +
  NoAxes() +
  coord_fixed(1)

```

### QC metrics

```{r, message = F, warning = F, fig.width = 8, fig.height = 8}

FeaturePlot(integrated_seurat,
            features = c("nCount_RNA", "nFeature_RNA", "percent.mt_RNA", "percent.ribo_RNA",
                         "S.Score", "G2M.Score"),
            ncol = 3,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

FeaturePlot(integrated_seurat,
            #layer = "RNA",
            slot = "data",
            features = c("CD4", "CD8A", "CD19", "PTPRC"),
            order = T,
            ncol = 2,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

```

### T cell markers

```{r, message = F, warning = F, fig.width = 12, fig.height = 12}

unique(canonical_markers$Bigger_gene_sets)

FeaturePlot(integrated_seurat,
            #layer = "RNA",
            slot = "data",
            features = canonical_markers[which(canonical_markers$Bigger_gene_sets == "Cytotoxic"),]$RNA,
            ncol = 4,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

FeaturePlot(integrated_seurat,
            #layer = "RNA",
            slot = "data",
            features = canonical_markers[which(canonical_markers$Bigger_gene_sets == "Dysfunction"),]$RNA,
            ncol = 4,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

FeaturePlot(integrated_seurat,
            #layer = "RNA",
            slot = "data",
            features = canonical_markers[which(canonical_markers$Bigger_gene_sets == "Memory"),]$RNA,
            ncol = 4,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

FeaturePlot(integrated_seurat,
            #layer = "RNA",
            slot = "data",
            features = canonical_markers[which(canonical_markers$Bigger_gene_sets == "Treg"),]$RNA,
            ncol = 4,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

```

### Top marker expression

```{r, fig.width = 12, fig.height=12}

DefaultAssay(integrated_seurat) <- "RNA"

# Top markers for each cluster
markers <- presto::wilcoxauc(integrated_seurat,
                             group_by = "cluster",
                             assay = "data",
                             seurat_assay = "RNA")

top_markers <- markers %>% group_by(group) %>% slice_max(order_by = auc, n = 2)

FeaturePlot(integrated_seurat,
            features = top_markers$feature,
            ncol = 6,
            reduction = reduction,
            raster = T,
            cols = c("gray89", "tomato3")) &
  coord_fixed(ratio = 1) &
  theme_bw() &
  NoLegend()

```

```{r, fig.width=10, fig.height=12}

top_markers <- markers %>%  group_by(group) %>% slice_max(order_by = auc, n = 5)

# seurat_object, plot_features, group_var, group_colors, column_title, km=5, row.order = NULL
dotplot_heatmap <- create_dotplot_heatmap(seurat_object = integrated_seurat,
                                          plot_features = unique(top_markers$feature),
                                          group_var = "cluster",
                                          group_colors = product_cluster_col,
                                          column_title = "",
                                          km = 5, row.order = NULL)

dotplot_heatmap <- create_dotplot_heatmap(seurat_object = integrated_seurat,
                                          plot_features = unique(canonical_markers$RNA),
                                          group_var = "cluster",
                                          group_colors = product_cluster_col,
                                          column_title = "",
                                          km = 5, row.order = NULL)

```

### Cell phenotypes by sample

```{r, message = F, warning = F, fig.width = 8, fig.height = 4}

# Create barplot
palette <- c("#edbe2a", "#A58AFF", "#53B400", "#F8766D", "#00B6EB", "#d1cfcc", "#FF0000",
             "#87f6a5", "#e812dd", "#777777")
names(palette) <- c(na.omit(unique(integrated_seurat$functional.cluster)), "NA")

integrated_seurat@meta.data$functional.cluster[is.na(integrated_seurat@meta.data$functional.cluster)] <- "NA"
integrated_seurat$functional.cluster <- factor(integrated_seurat$functional.cluster, levels = sort(unique(integrated_seurat$functional.cluster)))

# seurat_object, plot_var, group_var, group_levels, plot_levels, plot_colors, var_names, legend_title
barplot <- create_barplot(seurat_object = integrated_seurat,
                          plot_var = "functional.cluster",
                          group_var = "cluster",
                          group_levels = as.character(c(0, seq(1, max(unique(as.numeric(as.character(integrated_seurat$cluster))))))),
                          plot_levels = sort(unique(integrated_seurat$functional.cluster)),
                          plot_colors = palette,
                          var_names = c("% cells", "Cluster"),
                          legend_title = "T cell state")

barplot

integrated_seurat$annot <- mapvalues(x = integrated_seurat$cluster,
                                     from = cluster_annot$integratedSCTsnn_res.1,
                                     to = cluster_annot$cluster_name)

barplot <- create_barplot(seurat_object = integrated_seurat,
                          plot_var = "annot",
                          group_var = "Sample_Name",
                          group_levels = unique(integrated_seurat$Sample_Name),
                          plot_levels = sort(unique(integrated_seurat$annot)),
                          plot_colors = product_celltype_col,
                          var_names = c("% cells", "Sample"),
                          legend_title = "Cell type")

barplot

```

ARHGAP15 could be associated with naive state?
EBF1 required for B cell lineage commitment, for establishing B cell identity.
Not a lot of TIGIT expression, LAG3, or other dysfunction markers.

```{r, message = F, warning = F, fig.width = 8, fig.height = 7}

DimPlot(integrated_seurat,
        group.by = "annot",
        cols = product_celltype_col,
        reduction = reduction,
        label = T,
        label.box = T,
        label.size = 3,
        repel = T,
        raster = T,
        raster.dpi = c(1024, 1024),
        pt.size = 3) +
  ggtitle("Cell type") +
  theme_classic() +
  NoLegend() +
  NoAxes() +
  coord_fixed(1)

```

### Using sc-ImmuCC to annotate immune cells

```{r, results='hide', warning=F, message=F}

#count_data <- LayerData(integrated_seurat, assay = "RNA", layer = "counts")
#
#scImmuCC_Layered(count = count_data, Non_Immune = FALSE)
#
## Importing results
##scicc_labels <- read.csv("/home/hnatri/PD1_mm/docs/Layer1_scImmuCC_label.csv",
##                         row.names = "X")
#
#length(colnames(seurat_data))
#length(intersect(scicc_labels$barcodes, colnames(seurat_data)))
#
#seurat_data$scImmuCC_celltype <- mapvalues(x = colnames(seurat_data),
#                                           from = scicc_labels$barcodes,
#                                           to = scicc_labels$cell_type)
#
## Plotting
#DimPlot(seurat_data,
#        group.by = "scImmuCC_celltype",
#        reduction = "umap",
#        raster = T,
#        #cols = scImmuCC_celltype_col,
#        label = T) &
#  coord_fixed(ratio = 1) &
#  theme_bw() &
#  NoLegend() &
#  manuscript_theme

#saveRDS(seurat_data, "/tgen_labs/banovich/BCTCSF/PD1_mm_Seurat/PD1_mm_Seurat_merged.Rds")

```
