---
title: "19130: CAR T Production. Processing 10x data"
author: "heinin"
date: "2024-05-22"
output:
  workflowr::wflow_html:
    code_folding: hide
  html_notebook: default
editor_options:
  chunk_output_type: console
---

## Processing 10x data

### Libraries and environment variables
```{r, message = F, warning = F}

library(workflowr)
library(Seurat)
library(googlesheets4)
library(tidyverse)
library(plyr)
library(SoupX)
library(pheatbuilder)

setwd("/home/hnatri/19130_CAR_production/")
set.seed(1234)
options(future.globals.maxSize = 30000 * 1024^2)
reduction <- "umap"

source("/home/hnatri/SingleCellBestPractices/scripts/preprocessing_qc_module.R")

```

### Import 10x data, demultiplexing

The setup is two runs with the same samples:\
*Run1 (well 3A) F07344-GEX_F07352-CAR_F07348-FB, with a 50-60k pool load,\
*Run2 (well 3B) F07345-GEX_F07353-CAR_F07349-FB, with a 90-100k pool load

```{r, message = F, warning = F}

# Metadata
# https://docs.google.com/spreadsheets/d/1DstdK4vXjk7ZEeBalW_4CA6Vmd57BGENUx3iMDRLszk/edit?usp=sharing
gs4_deauth()
production_samples  <- gs4_get("https://docs.google.com/spreadsheets/d/1DstdK4vXjk7ZEeBalW_4CA6Vmd57BGENUx3iMDRLszk/edit?usp=sharing")
#sheet_names(production_samples)
production_samples <- read_sheet(production_samples, sheet = "Samples")

# Importing data. Two runs with the same samples.
# Concated GEX: /tgen_labs/banovich/SingleCell/CellRanger/7_1_0/Projects/BCTCSF/HT/BCTCSF_0130_1_PB_CONCAT_C1_X5HT2_F07344-F07345_CONCAT
path_list <- list("run1" = unique(production_samples[which(production_samples$Run=="run1"),]$CellRanger_path),
                  "run2" = unique(production_samples[which(production_samples$Run=="run2"),]$CellRanger_path))

# Cell hashing antibody names
hash_antibodies <- unique(production_samples$Hash_ID)

# Demultiplexing
seurat_list <- prep_seurat_list_multiplexed(metadata = production_samples,
                                            batch_ID = "Run",
                                            cellRanger_path = "CellRanger_path",
                                            cell_ID_prefix = "Sample_Name",
                                            CellHashing_Ab = "Hash_ID",
                                            Hash_Abs = hash_antibodies)
```

Singlets and doublets for each run.

```{r, message = F, warning = F, fig.width = 12, fig.height = 8}
# Singlets/doublets
table(seurat_list[["run1"]]$Hash_classification.global)
table(seurat_list[["run2"]]$Hash_classification.global)

p1 <- table(seurat_list[["run1"]]$Hash_classification.global) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash classification for run 1") +
  ylab("# cells") +
  NoLegend()

p2 <- table(seurat_list[["run1"]]$Hash_maxID) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash classification for run 1") +
  ylab("# cells") +
  NoLegend()

p3 <- table(seurat_list[["run2"]]$Hash_classification.global) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash classification for run 2") +
  ylab("# cells") +
  NoLegend()

p4 <- table(seurat_list[["run2"]]$Hash_maxID) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash classification for run 2") +
  ylab("# cells") +
  NoLegend()

(p1 + p2) / (p3 + p4)

```

Plotting HTO counts.

```{r, message = F, warning = F}

# Plotting HTO counts
hash_counts <- lapply(seurat_list, function(xx){
  hash <- LayerData(xx, assay = "Hash", layer = "counts")
  hash <- as.data.frame(t(hash))
  hash$hash_classification <- mapvalues(rownames(hash),
                             from = rownames(xx@meta.data),
                             to = xx@meta.data$Hash_maxID)
  
  hash
})

hash_antibodies <- gsub("_", "-", hash_antibodies)

```

Run 1

```{r, message = F, warning = F}

hash_counts[[1]] %>% pivot_longer(cols = hash_antibodies,
                                  values_to = "counts",
                                  names_to = "HashAB") %>%
  group_by(HashAB) %>%
  dplyr::summarize(sum = sum(counts)) %>%
  ungroup() %>%
  ggplot(aes(x = HashAB, y = sum, fill = HashAB)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash counts in run 1") +
  ylab("# cells")

hash_counts[[1]] %>% pivot_longer(cols = hash_antibodies,
                                  values_to = "counts",
                                  names_to = "HashAB") %>%
  group_by(hash_classification, HashAB) %>%
  dplyr::summarize(sum = sum(counts)) %>%
  ungroup() %>%
  pivot_wider(values_from = sum,
              names_from = HashAB) %>%
  pheat()

```

Run 2

```{r, message = F, warning = F}

hash_counts[[2]] %>% pivot_longer(cols = hash_antibodies,
                                  values_to = "counts",
                                  names_to = "HashAB") %>%
  group_by(HashAB) %>%
  dplyr::summarize(sum = sum(counts)) %>%
  ungroup() %>%
  ggplot(aes(x = HashAB, y = sum, fill = HashAB)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash counts in run 2") +
  ylab("# cells")

hash_counts[[2]] %>% pivot_longer(cols = hash_antibodies,
                                  values_to = "counts",
                                  names_to = "HashAB") %>%
  group_by(hash_classification, HashAB) %>%
  dplyr::summarize(sum = sum(counts)) %>%
  ungroup() %>%
  pivot_wider(values_from = sum,
              names_from = HashAB) %>%
  pheat()
```

Singlets and doublets in the combined data.

```{r, message = F, warning = F}
# Merging for plotting
seurat_merged <- merge(x = seurat_list[[1]], y = seurat_list[[2]])

# Singlets/doublets
table(seurat_merged$Hash_classification.global) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash classification for both runs") +
  ylab("# cells") +
  NoLegend()

table(seurat_merged$Hash_maxID) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Hash classification for both runs") +
  ylab("# cells") +
  NoLegend()

seurat_list <- lapply(seurat_list, function(xx){
  xx@meta.data$cellname <- colnames(xx)
  
  xx
})

```

Filtering to retain singlets

```{r, message = F, warning = F}

# Filtering to retain singlets
seurat_singlets_list <- lapply(seurat_list, function(xx){
  xx <- subset(xx, subset = Hash_classification.global == "Singlet")
  
  xx
})
names(seurat_singlets_list) <- names(seurat_list)

# Merging for plotting
seurat_merged <- merge(x = seurat_list[[1]], y = seurat_list[[2]])

# Plotting
table(seurat_merged$Sample_Name) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Sample") +
  ylab("# singets")

table(seurat_merged$Run) %>% as.data.frame() %>%
  ggplot(aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  xlab("Run") +
  ylab("# singlets")

```

### SoupX for ambient RNA

```{r, message = F, warning = F}

# Saving demultiplexed data in 10x format
for (i in 1:length(seurat_singlets_list)){
  print(paste0("Converting sample ", names(seurat_singlets_list[i])))
  obj.sub <- seurat_singlets_list[[i]] 
  
  counts <- LayerData(obj.sub, assay = "RNA", layer = "counts")
  DropletUtils::write10xCounts(path = paste0("/scratch/hnatri/CART/SoupX/demultiplexed_", names(seurat_singlets_list[i])), 
                               x = counts, 
                               barcodes = colnames(counts), # cell names
                               gene.id = rownames(counts), # Gene identifiers, one per row of X
                               type = "sparse", version="3", overwrite = T)
}

# Running SoupX
seurat_soupx_list <- sapply(names(seurat_singlets_list), function(xx){
  print(xx)

  # Read in count and droplet data
  # Converted demultiplexed counts
  message("Reading converted demultiplexed counts")
  d10x_toc <- Read10X(paste0("/scratch/hnatri/CART/SoupX/demultiplexed_", xx))
  
  #batch_id <- seurat_singlets_list[[xx]]$Run
  
  # Need to read in batch specific empty droplet file
  message("Reading empty droplet data")
  d10x_tod <- Read10X(paste0(path_list[[xx]], "/outs/raw_feature_bc_matrix/"))
  
  if(length(names(d10x_tod))>1){
    d10x_tod_gex <- d10x_tod[[1]]
  } else {
    d10x_tod_gex <- d10x_tod
  }
  
  colnames(d10x_tod_gex) <- paste0(xx, "_", colnames(d10x_tod_gex))
  
  # Some batches only have features that are expressed in at least one cell;
  # need to fix feature order
  #rownames(d10x_toc) <- gsub("_", "-", rownames(d10x_toc))
  #rownames(d10x_tod_gex) <- gsub("_", "-", rownames(d10x_tod_gex))
  d10x_tod_gex <- d10x_tod_gex[rownames(d10x_toc),]
  
  # Run SoupX
  sc <- SoupChannel(d10x_tod_gex, d10x_toc, calcSoupProfile = FALSE) 
  
  sc <- estimateSoup(sc)
  toc_seu <- CreateSeuratObject(d10x_toc)
  toc_seu <- SCTransform(toc_seu, vst.flavor = "v2")
  if(ncol(toc_seu)<50){
    toc_seu <- RunPCA(toc_seu, npcs = ncol(toc_seu)-1)
  } else{
    toc_seu <- RunPCA(toc_seu)
  }
  toc_seu <- RunUMAP(toc_seu, dims = 1:15)
  toc_seu <- FindNeighbors(toc_seu, dims = 1:15)
  toc_seu <- FindClusters(toc_seu, resolution = 1)

  # Add meta data to soupX object
  sc <- setClusters(sc, setNames(toc_seu$seurat_clusters, rownames(toc_seu@meta.data)))
  # Estimate contamination (automated method)
  message(paste0("Getting autoEstCont for: ", i))
  sc <- autoEstCont(sc, tfidfMin = 0.6, soupQuantile = 0.7, forceAccept = TRUE, doPlot = F) 
  out <- adjustCounts(sc)
  
  # Create Seurat object using corrected data
  d10x_seu <- CreateSeuratObject(out, assay = "SoupX_RNA")
  d10x_seu[["RNA"]] <- toc_seu@assays[["RNA"]]
  d10x_seu <- PercentageFeatureSet(d10x_seu, pattern = "^MT-", col.name = "percent.mt_RNA", assay = "RNA")
  d10x_seu <- PercentageFeatureSet(d10x_seu, pattern = "^RP[SL]|^MRP[SL]", col.name = "percent.ribo_RNA", assay = "RNA")
  d10x_seu <- PercentageFeatureSet(d10x_seu, pattern = "^MT-", col.name = "percent.mt_SoupX_RNA", assay = "SoupX_RNA")
  d10x_seu <- PercentageFeatureSet(d10x_seu, pattern = "^RP[SL]|^MRP[SL]", col.name = "percent.ribo_SoupX_RNA", assay = "SoupX_RNA")
  
  # Add sample metadata
  d10x_seu$Sample <- paste0(sapply(strsplit(colnames(d10x_seu), "_"), `[`, 1), "_",
                            sapply(strsplit(colnames(d10x_seu), "_"), `[`, 2))
  

  d10x_seu$Day <-paste0(sapply(strsplit(colnames(d10x_seu), "_"), `[`, 2))
  d10x_seu$Sample_Type <- paste0(sapply(strsplit(colnames(d10x_seu), "_"), `[`, 1))
  d10x_seu$cellname <- colnames(d10x_seu)
  d10x_seu$Run <- xx
  
  d10x_seu <- RenameCells(d10x_seu,
                          new.names = paste0(xx,"_", colnames(d10x_seu)))
  
  d10x_seu
})

names(seurat_soupx_list) <- names(seurat_singlets_list)

# Merge
seurat_merged <- merge(x = seurat_soupx_list[[1]], y = seurat_soupx_list[[2]])

#saveRDS(seurat_merged, "/scratch/hnatri/CART/SoupX/seurat_merged.rds")

```

### QC and filtering

```{r, message = F, warning = F}

par(mfrow=c(2,3))

# PLOTS 1 & 2: nCount vs. nFeature
smoothScatter(log2(seurat_merged$nCount_SoupX_RNA), log2(seurat_merged$nCount_SoupX_RNA),
              xlab = "log2(nCount_SoupX_RNA)", ylab = "log2(nFeature_SoupX_RNA)")

smoothScatter(seurat_merged$nCount_SoupX_RNA, seurat_merged$nCount_SoupX_RNA,
              xlab = "nCount_SoupX_RNA", ylab = "nFeature_SoupX_RNA")

# PLOTS 3 & 4: nCount vs. percent.mt_RNA
smoothScatter(seurat_merged$percent.mt_SoupX_RNA, log2(seurat_merged$nCount_SoupX_RNA),
              xlab = "% MT", ylab = "log2(nCount_SoupX_RNA)")

smoothScatter(seurat_merged$percent.mt_SoupX_RNA, seurat_merged$nCount_SoupX_RNA,
              xlab = "% MT", ylab = "nCount_SoupX_RNA")
abline(v = 10, h = 1000, 
       lty = "dashed", lwd = 1.25, col = "red")

# PLOTS 5 & 6: nFeature vs. percent.mt_RNA
smoothScatter(seurat_merged$percent.mt_SoupX_RNA, log2(seurat_merged$nFeature_SoupX_RNA),
              xlab = "% MT", ylab = "log2(nFeature_RNA)")

smoothScatter(seurat_merged$percent.mt_SoupX_RNA, seurat_merged$nFeature_SoupX_RNA,
              xlab = "% MT", ylab = "nFeature_RNA")
abline(v = 10, h = 500, 
       lty = "dashed", lwd = 1.25, col = "red")

seurat_filtered <- subset(seurat_merged, subset = percent.mt_SoupX_RNA < 10 &
                            nFeature_SoupX_RNA > 500 & nCount_SoupX_RNA >1000)

saveRDS(seurat_filtered, "/scratch/hnatri/CART/19130_CARprod_filtered.rds")

```


